---
title: "Hydrolysis 15 plates"
author: "Yuzi"
date: "25/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  results = 'asis',
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = 'center'
)
library(tidyverse)
```

# Notes (tidy the data)

The results of the following samples are marked as **NA** :

- 5 empty sapmles : 178, 101, 146, 107 and 119
- row H, column 3 to 12 in plate 15 : no samples
- col 5, 6, 9, 10 in plate 6 : technical issue (these sapmles were repeated in plate 15)
- the sample 49 in plate 6 : did not find this sample at that moment (have not been repeated in plate 15, might need to repeat it later)
- the sample 82* in plate 5 and 7 (did not distinguish from the sample 82)

# Scatter plot of the hydrolysis extent

```{r scatter-plot}
# import the dataset

data_15P_outlier_deleted <- read_csv("data/tidydata/data_15P_outlier_deleted.csv")

data_status <- data_15P_outlier_deleted %>% 
  mutate(status = case_when( # number of conditions (left), discription of the results in that condition
    Sample == "C+" ~ "Positive control", 
    Sample == "C-" ~ "Negative control",
    TRUE ~ "Sample"
  ))

ggplot(data = data_status, 
       aes(x = Time, 
           y = HE,
           group = Sample,
           color = status)) + # colored by the status (distinguish the control from the other samples)
  geom_point(size = 2, shape = 1) + # add the transparency
  scale_y_continuous(limits = c(0,100)) + ## set the range of the y axis
  scale_x_continuous(limits = c(0, 2000)) +
  ylab("Hydrolysis extent (%)") + ## change the label for the y axis
  xlab("Time (min)") + ## change the name of the x axis
  theme(legend.title = element_blank(),
        panel.grid = element_blank(),
        axis.line = element_line(colour = "black", size = 0.5)，
        panel.background = element_rect(fill = "white"),
        axis.ticks=element_line(
          colour="black",
          size=.5)) +
  labs(x = "Time (min)", y = "Hydrolysis extent (%)") +
  theme(axis.text.x = element_text(color="black", size=10), 
        axis.text.y = element_text(color="black", size=10)) +
  scale_color_discrete(labels = c("Negtive control", "Sample", "Positive control")) +
  theme(legend.key = element_blank(),
        legend.position = "bottom")


```


